### 2021-04-12 paper/ase21/analyzer.tex
[ASE21] Till 3.2

A modified $\ires$ program $\prog = (\getfunc, \getlab, \getnext)$ is consists
of three mappings from labels;  $\getfunc: \labset \rightarrow \funcset$
represents that labels are in which functions, $\getlab: \labset \rightarrow
\instset$ represents instructions attached to labels, and $\getnext: \labset
\rightarrow \labset$ denotes next labels where a label $\lab \in \labset$
denotes a program point.  function $\kwdef \; \f (\x^*, \y^*). \; \lab \in
\funcset$ consists of its name $\f$, normal parameters $\x^*$, optional
parameters $\y^*$, and a body label $\lab$.  Although global variables exist in
ECMAScript, we assume that only local variables exist in this paper for brevity
An instruction $\inst$ is a variable declaration, a function call, an assertion,
a branch, a return, or an reference update.  Each invocation of an abstract
algorithm in ECMAScript is compiled to a function call instruction with a newly
introduced temporal variable.  We represent loops using branch instructions with
cyclic pointing of labels in the next mapping $\getnext$.  A reference $\refer$
is a variable $\x$ or a field access $\refer \kwsl \expr \kwsr$.  An expression
$\expr$ is a record, a list, a type check, an existence check, a binary
operation, a unary operation, a reference, a constant, or a primitive, which is
either $\undefval$, $\nullval$, a Boolean $\bool$, a Number $\num$, a BigInt
$\bigint$, a String $\str$, or a Symbol $\symb$.

@@

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
described in Figure~\ref{fig:subtype}; the directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$ and the relation is reflexive and transitive.  The
subtype relation is dependent on the syntax and defined record types in
ECMAScript.  We extract subtypes for production types and manually model
subtypes between record types.  For example, consider the syntax-directed
abstract algorithm at the right bottom in Figure~\ref{fig:example}.  The
non-terminal \textit{BindingElement} is the unique token in a alternative of
\textit{FormalParameter} production thus we automatically extract the subtype
relation: \textit{BindingElement} $\subtype$ \textit{FormalParameter}.  Based on
this subtype relation, a type check expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has a type $\ty'$ satisfying $\ty' \subtype \ty$.

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{\textwidth}
    \[
      \begin{array}{r@{~}c@{~}ll}
        \asemi{\kwlet \; \x = \expr}
        (\lab, \tys, \aenv, \_) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\x \}, \emp) &
        \text{where} \;
        \aenv_\x = \aenv[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\x = \kwrl \expr_0 \; \expr_1 \cdots \expr_n \kwrr}
        (\lab, \tys, \aenv, \_) &=&
        \left(
          \begin{array}{@{}l@{}}
            \{ (\lab_\func, \tys_a) \mapsto \aenv_{\func, \tys_a} \mid \func \in
              F \wedge \tys_a \in T_a
            \},\\
            \{ (\func, \tys_a) \mapsto \{ (\getnext(\lab), \tys, \x) \} \mid
              \func \in F \wedge \tys_a \in T_a
            \}\\
          \end{array}
        \right) &
        \text{where} \;
        \left\{
          \begin{array}{@{}l@{}}
            \aty_i = \aseme{\expr_i}(\aenv) \wedge T_a = \{ \upcasts([ \ty_1,
            \cdots, \ty_n ] \mid \ty_i \in \aty_i \} \wedge\\
            F = \{ \func \in \aty_0 \mid \func = \kwdef \; \_ (\p_1, \cdots,
            \p_m). \; \lab_\func \} \wedge\\
            \aenv_{\func, \tys_a} = \aenv[\p_1 \mapsto \{ \tys_a[1] \}, \cdots,
            \p_m \mapsto \{ \tys_a[m] \}]\\
          \end{array}
        \right.\\

        \asemi{\kwassert \; \expr}
        (\lab, \tys, \aenv, \_) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\mt \}, \emp) &
        \text{where} \; \aenv_\mt = \refine(\aenv, \expr, \true)\\

        \asemi{\kwif \; \expr \; \lab_\mt \; \lab_\mf}
        (\lab, \tys, \aenv, \_) &=&
        (\{ (\lab_\mt, \tys) \mapsto \aenv_\mt, (\lab_\mf, \tys) \mapsto
        \aenv_\mf \}, \emp) &
        \text{where} \;
        \aenv_\mt = \refine(\aenv, \expr, \true) \wedge
        \aenv_\mf = \refine(\aenv, \expr, \false)\\

        \asemi{\kwreturn \; \expr}
        (\lab, \tys, \aenv, \aelem) &=&
        (\{ (\lab_r, \tys_r) \mapsto \aenv_r \mid (\lab_r, \tys_r, \x) \in
        \retp(\getfunc(\lab), \tys)\}, \emp) &
        \text{where} \;
        \aelem = (\rmap, \retp) \wedge
        \aenv_r = \rmap(\lab_r, \tys_r)[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\x = \expr}
        (\lab, \tys, \aenv, \_) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\x \}, \emp) &
        \text{where} \;
        \aenv_\x = \aenv[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\refer \kwsl \expr_0 \kwsr = \expr_1}
        (\lab, \tys, \aenv, \_) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv \}, \emp)
      \end{array}
    \]
    \caption{Instructions with $\asemi{\inst}: (\labset \times \tyset^* \times
    \aenvset \times \adom) \rightarrow \adom$}
  \end{subfigure}
  \begin{subfigure}[b]{0.48\textwidth}
    \[
      \asemr{\x}(\aenv) = \aenv(\x)
      \qquad
      \asemr{\refer \kwsl \expr \kwsr}(\aenv) = \{ \ty[\str] \mid \ty \in
      \asemr{\refer}(\aenv) \wedge \str \in \aseme{\expr}(\aenv) \}
    \]
    \caption{References with $\asemr{\refer}: \aenvset \rightarrow \atyset$}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \[
      \begin{array}{r@{~}c@{~}lr@{~}c@{~}lr@{~}c@{~}l}
        \aseme{\tname \; \kwcl \fields \kwcr}(\aenv)
        &=& \multicolumn{7}{@{}l}{
          \left\{
            \begin{array}{ll}
              \{ \tnormal(\ty) \mid \ty \in \aseme{\fields(\code{Value})}(\aenv)
              \} & \text{if} \; \tname = \text{Completion} \wedge
              \fields(\code{Type}) = \nconst{normal}\\

@@

        \\
      \end{array}
    \]
    \caption{Expressions with $\aseme{\expr}: \aenvset \rightarrow \atyset$}
  \end{subfigure}
  \caption{Abstract semantics of the modified $\ires$ for a program $\prog =
  (\getfunc, \getlab, \getnext)$.}
  \vspace*{-1.5em}
  \label{fig:abs-sem}
\end{figure*}

@@

    \text{Abstract Types}
    &\aty&\in&\atyset&=&\partsof{\tyset}\\
  \end{array}
\]
An abstract state $\aelem \in \adom$ is pair of a return map and a return point
mapping.  A result map $\rmap \in \rmapset$ represents an abstract environment
for each flow-sensitive and type-sensitive view, and a return point mapping
$\retp \in \retpset$ represents possible return points of each function with
a type-sensitive context.  An abstract environment $\aenv \in \aenvset$
represents possible types of each variable, and $\aenv(\x) = \{ \tabsent \}$
when $\x$ is not defined in $\aenv$.  An abstract type $\aty \in \atyset$ is a
set of extended types.

We define the abstract semantics of the modified $\ires$ as described in
Figure~\ref{fig:abs-sem} for a program $\prog = (\getfunc, \getlab,
\getnext)$.

@@

>>>>

A modified $\ires$ program $\prog = (\getfunc, \getinst, \getnext)$ is consists
of three mappings from labels;  $\getfunc: \labset \rightarrow \funcset$
represents that labels are in which functions, $\getinst: \labset \rightarrow
\instset$ represents instructions attached to labels, and $\getnext: \labset
\rightarrow \labset$ denotes next labels where a label $\lab \in \labset$
denotes a program point.  function $\kwdef \; \f (\x^*, \y^*). \; \lab \in
\funcset$ consists of its name $\f$, normal parameters $\x^*$, optional
parameters $\y^*$, and a body label $\lab$.  Although global variables exist in
ECMAScript, we assume that only local variables exist in this paper for brevity
An instruction $\inst$ is a variable declaration, a function call, an assertion,
a branch, a return, or an reference update.  Each invocation of an abstract
algorithm in ECMAScript is compiled to a function call instruction with a newly
introduced temporal variable.  We represent loops using branch instructions with
cyclic pointing of labels in the next mapping $\getnext$.  A reference $\refer$
is a variable $\x$ or a field access $\refer \kwsl \expr \kwsr$.  An expression
$\expr$ is a record, a list, a type check, an existence check, a binary
operation, a unary operation, a reference, a constant, or a primitive, which is
either $\undefval$, $\nullval$, a Boolean $\bool$, a Number $\num$, a BigInt
$\bigint$, a String $\str$, or a Symbol $\symb$.

@@

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
described in Figure~\ref{fig:subtype}; the directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$ and the relation is reflexive and transitive.  The
subtype relation is dependent on the syntax and defined record types in
ECMAScript.  We extract subtypes for production types and manually model
subtypes between record types.  For example, consider the syntax-directed
abstract algorithm at the lower-right in Figure~\ref{fig:example}.  The
non-terminal \textit{BindingElement} is the unique token in a alternative of
\textit{FormalParameter} production thus we automatically extract the subtype
relation: \textit{BindingElement} $\subtype$ \textit{FormalParameter}.  Based on
this subtype relation, a type check expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has a type $\ty'$ satisfying $\ty' \subtype \ty$.

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{\textwidth}
    \[
      \begin{array}{r@{~}c@{~}ll}
        \asemi{\kwlet \; \x = \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\x \}, \emp) &
        \text{where} \;
        \aenv_\x = \aenv[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\x = \kwrl \expr_0 \; \expr_1 \cdots \expr_n \kwrr}
        (\lab, \tys)(\aelem) &=&
        \left(
          \begin{array}{@{}l@{}}
            \{ (\lab_\func, \tys_a) \mapsto \aenv_{\func, \tys_a} \mid \func \in
              F \wedge \tys_a \in T_a
            \},\\
            \{ (\func, \tys_a) \mapsto \{ (\getnext(\lab), \tys, \x) \} \mid
              \func \in F \wedge \tys_a \in T_a
            \}\\
          \end{array}
        \right) &
        \text{where} \;
        \left\{
          \begin{array}{@{}l@{}}
            \aty_i = \aseme{\expr_i}(\aenv) \wedge T_a = \{ \upcasts([ \ty_1,
            \cdots, \ty_n ] \mid \ty_i \in \aty_i \} \wedge\\
            F = \{ \func \in \aty_0 \mid \func = \kwdef \; \_ (\p_1, \cdots,
            \p_m). \; \lab_\func \} \wedge\\
            \aenv_{\func, \tys_a} = \aenv[\p_1 \mapsto \{ \tys_a[1] \}, \cdots,
            \p_m \mapsto \{ \tys_a[m] \}]\\
          \end{array}
        \right.\\

        \asemi{\kwassert \; \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\mt \}, \emp) &
        \text{where} \; \aenv_\mt = \refine(\aenv, \expr, \true)\\

        \asemi{\kwif \; \expr \; \lab_\mt \; \lab_\mf}
        (\lab, \tys)(\aelem) &=&
        (\{ (\lab_\mt, \tys) \mapsto \aenv_\mt, (\lab_\mf, \tys) \mapsto
        \aenv_\mf \}, \emp) &
        \text{where} \;
        \aenv_\mt = \refine(\aenv, \expr, \true) \wedge
        \aenv_\mf = \refine(\aenv, \expr, \false)\\

        \asemi{\kwreturn \; \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\lab_r, \tys_r) \mapsto \aenv_r \mid (\lab_r, \tys_r, \x) \in
        \retp(\getfunc(\lab), \tys)\}, \emp) &
        \text{where} \;
        \aelem = (\rmap, \retp) \wedge
        \aenv_r = \rmap(\lab_r, \tys_r)[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\x = \expr}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv_\x \}, \emp) &
        \text{where} \;
        \aenv_\x = \aenv[\x \mapsto \aty] \wedge
        \aty = \aseme{\expr}(\aenv)\\

        \asemi{\refer \kwsl \expr_0 \kwsr = \expr_1}
        (\lab, \tys)(\aelem) &=&
        (\{ (\getnext(\lab), \tys) \mapsto \aenv \}, \emp)
      \end{array}
    \]
    \caption{Instructions with $\asemi{\inst}: (\labset \times \tyset^*)
    \rightarrow \adom \rightarrow \adom$ where $\aenv = \aelem(\lab, \tys)$.}
  \end{subfigure}
  \begin{subfigure}[b]{0.48\textwidth}
    \[
      \asemr{\x}(\aenv) = \aenv(\x)
      \qquad
      \asemr{\refer \kwsl \expr \kwsr}(\aenv) = \{ \ty[\str] \mid \ty \in
      \asemr{\refer}(\aenv) \wedge \str \in \aseme{\expr}(\aenv) \}
    \]
    \caption{References with $\asemr{\refer}: \aenvset \rightarrow \atyset$.}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \[
      \begin{array}{r@{~}c@{~}lr@{~}c@{~}lr@{~}c@{~}l}
        \aseme{\tname \; \kwcl \fields \kwcr}(\aenv)
        &=& \multicolumn{7}{@{}l}{
          \left\{
            \begin{array}{ll}
              \{ \tnormal(\ty) \mid \ty \in \aseme{\fields(\code{Value})}(\aenv)
              \} & \text{if} \; \tname = \text{Completion} \wedge
              \fields(\code{Type}) = \nconst{normal}\\

@@

        \\
      \end{array}
    \]
    \caption{Expressions with $\aseme{\expr}: \aenvset \rightarrow \atyset$.}
  \end{subfigure}
  \caption{Abstract semantics of the modified $\ires$ for a program $\prog =
  (\getfunc, \getinst, \getnext)$.}
  \vspace*{-1.5em}
  \label{fig:abs-sem}
\end{figure*}

@@

    \text{Abstract Types}
    &\aty&\in&\atyset&=&\partsof{\tyset}\\
  \end{array}
\]
An abstract state $\aelem \in \adom$ is pair of a return map and a return point
mapping.  A result map $\rmap \in \rmapset$ represents an abstract environment
for each flow-sensitive and type-sensitive view, and a return point mapping
$\retp \in \retpset$ represents possible return points of each function with
a type-sensitive context.  An abstract environment $\aenv \in \aenvset$
represents possible types of each variable, and $\aenv(\x) = \{ \tabsent \}$
when $\x$ is not defined in $\aenv$.  An abstract type $\aty \in \atyset$ is a
set of types.  While the join operator $\join$ and the partial order $\order$
for abstract states are defined in a point-wise manner, the operators for types
are defined with a normalization function $\norm$ because of their subtype
relations as follows:
\[
  \begin{array}{l}
    \aty_0 \join \aty_1 = \norm(\aty_0 \cup \aty_1)\\
    \aty_0 \order \aty_1 \Leftrightarrow \forall \ty_0 \in \aty_0. \; \exists
    \ty_1 \in \norm(\aty_1). \; \text{s.t.} \; \ty_0 \subtype \ty_1\\
  \end{array}
\]
where $\norm(\aty) = \{ \ty \mid \ty \in \aty_e \wedge \not\exists \ty' \in
\aty_e. \; \text{s.t.} \; \ty \neq \ty' \wedge \ty \subtype \ty' \}$.

Based on abstract states, we define the abstract semantics of instructions,
references, and expressions in the modified $\ires$ described in
Figure~\ref{fig:abs-sem}.  To prevent the explosion of the number of type-sensitive
views, we upcast the type of each argument before function calls with the
following function:
\[
  \upcast(\ty) = \left\{
    \begin{array}{ll}
      \tnormal(\upcast(\ty')) & \text{if} \; \ty = \tnormal(\ty')\\
      \clist{\upcast(\ty')} & \text{if} \; \ty = \clist{\ty'}\\
      \tstr & \text{if} \; \ty = \str\\
      \tbool & \text{if} \; \ty = \bool\\
      \ty & \text{otherwise}\\
    \end{array}
  \right.
\]
and $\upcasts$ denotes a point-wise extension of $\upcast$ for type sequences.
For conditions, we apply the $\refine$ function that prunes out infeasible parts
in abstract states based on given conditions to increase analysis precision.  We
explain its detail in Section~\ref{sec:refine}.  Then, we define the abstract
semantics $\asem{\prog}$ of a program $\prog = (\getfunc, \getinst, \getnext)$
as the least fixpoint of abstract transfer $\atransfer: \adom \rightarrow \adom$
as follows:
\[
  \begin{array}{r@{~}c@{~}l}
    \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
    \atransfer(\aelem) &=& \aelem \join \left(
      \underset{(\lab, \tys) \in \domain{\rmap}}{\bigjoin}
      \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
    \right)\\
  \end{array}
\]
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in the Section~\ref{sec:overview}, $\iaelem$ contains all
syntax-directed algorithms having no additional parameters and built-in
algorithms with appropriate abstract environments.  For syntax-directed
algorithms, we construct its abstract environment containing the variable
$\this$ with a production type and other variables for non-terminals.  For
example, the upper-left syntax-directed algorithm in Figure~\ref{fig:example} is
initialized with the following abstract environment:
\[
  \left\{
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \text{AssignmentExpression} \},\\
      \mcode{LeftHandSideExpression} &\mapsto& \{ \text{LeftHandSideExpression} \},\\
      \mcode{AssignmentExpression} &\mapsto& \{ \text{AssignmentExpression} \}
    \end{array}
  \right.
\]
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ at the upper-right in Figure~\ref{fig:example}:
\[
  \left\{
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \},
      \args \mapsto \{ \clist{\tjs} \},
      \x \mapsto \tjs\\
      \NewTarget &\mapsto& \{ \text{Object}, \undefval \},\\
    \end{array}
  \right.
\]

@@

\todo
